---
output: 
  html_document:
    theme: yeti
    css: styles.css
    includes:
      before_body: header.html
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("MASS")
library("dplyr")
library("ggplot2")
library("grid")
library("stringr")
library("reshape2")
library("car")
library("PerformanceAnalytics")


irs_2011_2015 <- readRDS('../../r-objects/irs_2011_2015.rds')
education_df_plus_3 <- readRDS('../../r-objects/education_df_plus_3.rds')
zip_codes_df <- readRDS('../../r-objects/zip_irs_education_df.rds')
demo_act <- readRDS('../../r-objects/zip_irs_education_df.rds')[, c("Pct_Black", "Pct_Hispanic", "ACT_Composite", "county", "income_per_tax_return")]

```

```{r, echo=FALSE}
df_sort <- irs_2011_2015 %>% 
  filter(income_per_tax_return >= 8) %>% 
  select(county, year, income_per_tax_return)

```

<div class="row question-row">
  <blockquote class="blockquote">
    Q1: What is the income distribution amongst TN counties?
  </blockquote>
  <div class="col-md-12">
```{r, echo=FALSE}

h <- hist(df_sort$income_per_tax_return, breaks=10, plot=FALSE)
cuts <- cut(h$breaks, c(-Inf, 9, 19, 23, 25, Inf))
plot(
  h, 
  col=c('#062F4F', '#6D7993', '#B82601', '#6D7993', '#B82601')[cuts],
  main = '',
  xlab = 'Household Income Range * 1000',
  ylab = 'County Frequency'
)

```
  </div>
  <p class="lead"><em>** The majority of the county's household income falls between $2000 - $17000.</em></p>
</div>




<div class="row question-row">
  <blockquote class="blockquote">
    Q2: Is there any change in income over the years?
  </blockquote>
  <div class="col-md-12">
```{r irs_2011_2015, echo = FALSE}
df_sort <- irs_2011_2015 %>% 
  filter(income_per_tax_return >= 8) %>% 
  select(
    county, year, income_per_tax_return
  )

df_sort %>% 
  ggplot(
    aes(
      x = year,
      y = income_per_tax_return, 
      group = county, 
      color = county
    )
  ) + geom_line(size = 1.5) + 
  labs(y = "Income Per Tax Return * 1000", x = "Years")

```
  </div>
  <p class="lead"><em>** The majority of counties maintain their rank amongst other counties.</em></p>
</div>




```{r, echo=FALSE}
ggplot_num_returns <- function(df, range, y.label="") {
  df_sort_1 <- irs_2011_2015 %>% 
    dplyr::filter(sum_total_income_returns >= range) %>% 
    dplyr::select(
      county, year, sum_total_income_returns
    )
  
  df_sort_1 %>% 
    ggplot(
      aes(
        x = year,
        y = sum_total_income_returns, 
        group = county,
        color = county
      )
    ) + 
    geom_line(size = 1.5) + 
    labs(y = y.label, x = "") +
    theme(axis.text.x = element_text(
        face = 'bold', 
        size = 14
      )
    ) +
    theme(axis.text.y = element_text(
        face = 'bold', 
        size = 14
      )
    ) +
    theme(axis.title.y = element_text(
        size = 20
      )
    ) +
    scale_y_continuous(labels = scales::comma) +
    scale_color_hue(l = 60, c = 50)
}

```

<div class="row question-row">
  <blockquote class="blockquote">
    Q3: Which counties have highest number of tax filers per year?
  </blockquote>
  <div class="col-md-6">
```{r, echo=FALSE}
ggplot_num_returns(irs_2011_2015, 20000, y.label = "Num of Tax Payers Per Year")
```
  </div>
  <div class="col-md-6">
```{r, echo=FALSE}
ggplot_num_returns(irs_2011_2015, 200000)
```
  </div>
  <p class="lead"><em>** <strong>Davidson County</strong> is having a consistent increase in the number of tax filers, while <strong>Shelby County</strong> is decreasing.</em></p>
</div>



```{r, echo=FALSE}
create_hist <- function(df, col.name, col.breaks, col.colors, main, y.lab='') {
  h <- hist(df[[col.name]], breaks=10, plot=FALSE)
  cuts <- cut(h$breaks, col.breaks)
  plot(
    h, 
    col=col.colors[cuts],
    main = main,
    ylab = y.lab,
    xlab = ''
  )
}
```


<div class="row question-row">
  <blockquote class="blockquote">
    Q4: For 2014, what is the overall situation within our education dataset regarding student graduation, dropout, and suspension?
  </blockquote>
  <div class="col-md-12">
```{r, echo=FALSE}
par(mfrow=c(1, 3))
education_df_plus_3 %>% 
  create_hist('Graduation', c(-Inf, 82, 91, 95, Inf), c('#B82601', '#6D7993', '#062F4F', '#6D7993'), 'Graduation', y.lab = 'County Frequency')

education_df_plus_3 %>% 
  create_hist('Dropout', c(-Inf, 1, 3, 4, 11, Inf), c('#6D7993', '#062F4F', '#6D7993', '#6D7993', '#B82601'), 'Dropout')

education_df_plus_3 %>% 
  create_hist('Pct_Suspended', c(-Inf, 1, 6, 11, 13, Inf), c('#062F4F', '#6D7993', '#B82601', '#6D7993', '#B82601'), 'Pct Suspended')

```
  </div>

  <div class="col-md-12">
    
* Graduation rate is highest in Meigs, lowest in <strong><em>Shelby</em></strong>.
* Dropout rate is highest in <strong><em>Shelby</em></strong> county, while kids are studious in Pickett and Meigs counties.
* Teachers seem to be <em>very</em> strict in <strong><em>Shelby</em></strong> county, since it holds the highest student suspension rate.

  </div>
</div>




<div class="row question-row">
  <blockquote class="blockquote">
    Q5: For 2014, what is the overall situation within our education dataset regarding student performance?
  </blockquote>
  <div class="col-md-12">
```{r, echo=FALSE}
par(mfrow=c(2, 2))
education_df_plus_3 %>% 
  create_hist('ACT_Composite', c(-Inf, 17, 18, 20, Inf), c('#6D7993', '#062F4F', '#6D7993', '#B82601'), 'ACT_Composite', y.lab = 'County Frequency')

education_df_plus_3 %>% 
  create_hist('mean_math', c(-Inf, 19, 39, 69, 79, 89, Inf), c('#6D7993', '#B82601', '#6D7993', '#062F4F', '#6D7993', '#B82601'), 'Avg Math')

education_df_plus_3 %>% 
  create_hist('mean_science', c(-Inf, 39, 69, 79, 89, Inf), c('#B82601', '#6D7993', '#062F4F', '#6D7993', '#B82601'), 'Avg Science', y.lab = 'County Frequency')

education_df_plus_3 %>% 
  create_hist('mean_literature', c(-Inf, 29, 44, 49, 69, Inf), c('#B82601', '#6D7993', '#062F4F', '#6D7993', '#B82601'), 'Avg Literature')

```
  </div>

* <strong><em>Williamson</em></strong> county is showing highest ACT scores, average mathematics, literature, and second highest in science scores.
* The lowest ACT scores are from Lake and Fayette counties
* The lowest mathematics scores are from Cheatham and Putnam counties.
* <strong><em>Fentress</em></strong> takes the lead for the highest science scores at 100%, with Lake being lowest, and <strong><em>Shelby</em></strong> 3rd lowest.
* The lowest literature scores are from Hancock county.

</div>



<div class="row question-row">
  <blockquote class="blockquote">
    Q6: For 2014, what is the average ACT Scores per region within TN?
  </blockquote>
  <div class="md-col-12">
```{r, echo=FALSE, message=FALSE, warnings=FALSE}

# Subset the education data to only contain the act composite, math, science, and lit cols
edu_act_vs_core_region <- education_df_plus_3[, c("ACT_Composite", "CORE_region")]

# Melt the df to be indexed by Act composite
md_act_core <- melt(edu_act_vs_core_region, id="ACT_Composite")

md_act_core_labels <- c("East TN", "First TN", "Mid Cumberland", "Northwest", "South Central", "Southeast", "Southwest/Memphis", "Upper Cumberland")

md_act_core %>% 
  ggplot(aes(
    value,
    ACT_Composite
  )) + 
  geom_boxplot() + 
  geom_dotplot(
    binaxis='y', 
    stackdir='center', 
    dotsize = .5, 
    fill="red"
  ) +
  theme(
    axis.text.x = element_text(angle=55, vjust=0.6),
    panel.background = element_rect(
      fill = "#6D7993",
      colour = "#6D7993",
      size = 0.5, 
      linetype = "solid"
    ),
    
  ) + 
  labs(
    title="ACT Scores per CORE Region", 
    subtitle="Each dot represents 1 row in source data",
    x="",
    y="ACT Score"
  ) +
  scale_x_discrete(labels = md_act_core_labels)

```
  </div>

<p class="lead">Mid Cumberland show the highest average overall, which also includes Williamson county.</p>

</div>






<div class="row question-row">
  <blockquote class="blockquote">
    Q7: How do dropout, graduation, and ACT scores compare to income scales?
  </blockquote>
  <div class="col-md-12>

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=10}

zip_codes_df$income_category <- lapply(zip_codes_df$income_category, function(cat) {
  str_split(cat, ' Income')[[1]][[1]]
})

zip_codes_df$income_category <- factor(
  zip_codes_df$income_category, 
  levels = c('Low', 'Medium', 'High', 'Very High'),
  ordered = TRUE
)

# par(mfrow=c(1, 3))


plot1 <- zip_codes_df %>% 
  ggplot(aes(
    income_category,
    Dropout
  )) + 
  geom_boxplot(
    color = c("#062F4F", "#6D7993", "#6D7993", "#B82601")
  ) + 
  geom_dotplot(
    binaxis='y', 
    stackdir='center', 
    dotsize = .5, 
    fill="black"
  ) +
  theme(
    axis.text.x = element_blank(),
    panel.background = element_rect(
      size = 0.5, 
      linetype = "solid"
    )
  ) + 
  labs(
    title="Dropout Rates per Income Bucket", 
    subtitle="Each dot represents 1 row in source data",
    x="",
    y="Dropout"
  )

plot2 <- zip_codes_df %>% 
  ggplot(aes(
    income_category,
    Graduation
  )) + 
  geom_boxplot(
    color = c("#062F4F", "#6D7993", "#6D7993", "#B82601")
  ) + 
  geom_dotplot(
    binaxis='y', 
    stackdir='center', 
    dotsize = .5, 
    fill="black"
  ) +
  theme(
    axis.text.x = element_blank(),
    panel.background = element_rect(
      size = 0.5, 
      linetype = "solid"
    )
  ) + 
  labs(
    title="Graduation Rates per Income Bucket", 
    subtitle="Each dot represents 1 row in source data",
    x="",
    y="Graduation"
  )

plot3 <- zip_codes_df %>% 
  ggplot(aes(
    income_category,
    ACT_Composite
  )) + 
  geom_boxplot(
    color = c("#062F4F", "#6D7993", "#6D7993", "#B82601")
  ) + 
  geom_dotplot(
    binaxis='y', 
    stackdir='center', 
    dotsize = .5, 
    fill="black"
  ) +
  theme(
    axis.text.x = element_text(angle=55, vjust=0.6),
    panel.background = element_rect(
      size = 0.5, 
      linetype = "solid"
    )
  ) + 
  labs(
    title="ACT Scores per Income Bucket", 
    subtitle="Each dot represents 1 row in source data",
    x="",
    y="ACT Composite"
  )

grid.newpage()
grid.draw(
  rbind(ggplotGrob(plot1), ggplotGrob(plot2), ggplotGrob(plot3), size = "last")
)

```
  </div>
  <p class="lead">We notice that the medium and high income groups rank highest across all three variables. They have the highest graduation rate, lowest dropout rate, and overall highest ACT Scores.</p>
</div>

<div class="row question-row">
  <blockquote class="blockquote">
    Act Scores for the Black and Hispanic populations within TN counties, with size based on Income.
  </blockquote>
  
  <div class="col-md-12">
```{r, echo=FALSE, warning=FALSE, message=FALSE}
# Melt the df to be indexed by Act compositea
md_demo_act <- melt(demo_act, id=c("ACT_Composite", "county", "income_per_tax_return"))

md_demo_act %>% 
  ggplot(aes(
    x = value,
    y = ACT_Composite,
    color = variable,
    size = income_per_tax_return
  )) + 
  geom_point() +
  scale_colour_manual(values=c('#B82601', '#062F4F')) +
  geom_smooth(method='lm') + 
  labs(
    title="Avg ACT Scores By Demographics With Size Based on Income Levels",
    y="ACT Score",
    x = "Population Percentage"
  ) +
  theme(axis.text.y = element_text(
      face = 'bold'
    )
  ) +
  theme(axis.title.y = element_text(
      face = 'bold'
    )
  )

```
  </div>
</div>



<div class="row question-row">
  <blockquote class="blockquote">
    Performance Analytics Summary
  </blockquote>
  
  <div class="col-md-12">
```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.height=10,fig.width=10}
cor_df <- zip_codes_df[ , c(2:7,9:11,15)]
chart.Correlation(cor_df, histogram = T, pch = 19)
```
  </div>
</div>




