---
output: 
  html_document:
    theme: yeti
    css: styles.css
    includes:
      before_body: header.html
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("MASS")
library("dplyr")
library("ggplot2")
library("ggalt")
library("grid")
library("stringr")
library("reshape2")
library("car")
library("PerformanceAnalytics")


irs_2011_2015 <- readRDS('../../r-objects/irs_2011_2015.rds')
education_df_plus_3 <- readRDS('../../r-objects/education_df_plus_3.rds')
zip_codes_df <- readRDS('../../r-objects/zip_irs_education_df.rds')
demo_act <- readRDS('../../r-objects/zip_irs_education_df.rds')[, c("Pct_Black", "Pct_Hispanic", "ACT_Composite", "county", "income_per_tax_return")]

```

<div class="row question-row">
  <blockquote class="blockquote">
    Act Scores Black and Hispanic Populations
  </blockquote>
  
  <div class="col-md-12">
```{r, echo=FALSE, warning=FALSE, message=FALSE}
# Melt the df to be indexed by Act compositea
md_demo_act <- melt(demo_act, id=c("ACT_Composite", "county", "income_per_tax_return"))

md_demo_act_shelby_williamson <- md_demo_act %>% 
  dplyr::filter(county == 'williamson' | county == 'shelby')

md_demo_act %>% 
  ggplot(aes(
    x = value,
    y = ACT_Composite,
    color = variable,
    size = income_per_tax_return
  )) + 
  geom_point() +
  geom_point(
    aes(value, ACT_Composite),
    data = md_demo_act_shelby_williamson,
    color = 'purple'
  ) +
  scale_colour_manual(values=c('#B82601', '#062F4F')) +
  geom_smooth(method='lm') + 
  labs(
    title="Avg ACT Scores By Demographics With Size Based on Income Levels",
    y="ACT Score",
    x = "Population Percentage"
  ) +
  theme(axis.text.y = element_text(
      face = 'bold'
    )
  ) +
  theme(axis.title.y = element_text(
      face = 'bold'
    )
  )

```
  </div>
<div class="description">
  <p>This graph really helps paint a better picture regarding ACT vs Income, while also incorporating racial demographics as a factor.</p>
  <p>With color relating to race, and size shown by income levels, you can pretty clearly see that there is not a consistent trend regarding income. There are some counties with a high income level doing <em>FANTASTIC</em> on the ACT (such as <strong><em>Williamson</em></strong> at <strong><em>23.8</em></strong>), as well as some mid level counties. Similarly, there are some higher incomes doing <em>POORLY</em> on the ACT.</p>
  <p>However, there does seem to be at least a small, but not insignificant, correlation between the racial population and education levels within counties. As this graph portrays, counties with a higher black population tend to do worse on the ACT vs the hispanic population, with the highest being.. <strong>><em>Shelby</em></strong> at <strong><em>78.4%</em></strong> black population.</p>
</div>
</div>